Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets

### 2021-06-04 main.tex
pass section trees and parsing

Let $\Omega$ be a countable ranked alphabet, such that 
every symbol $a \in \Omega$ has a rank 
$\rank(a) \in [0..M]$ where $M$ is a fixed natural number.
We denote by $\Omega_k$ the subset of all symbols $a$ of $\Omega$
with $\rank(a) = k$, where $0 \leq k \leq M$, 
and $\Omega_{>0} = \Omega \setminus \Omega_0$.
%
\noindent 
The free $\Omega$-algebra of finite, ordered, 
$\Omega$-labeled trees is denoted by $\T(\Omega)$.
It is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
and for all $1 \leq k \leq M$, all $a \in \Omega_k$, 
and all $t_1, \ldots, t_k \in \T(\Omega)$, $a(t_1, \ldots, t_k) \in \T(\Omega)$.
%
% tree = single node (leave) labeled with a symbol of $a \in \Omegai$
% (such a tree is simply denoted by $a$)
% or the composition, denoted by $b(t_1,\ldots, t_n$) of a node labeled with $b$
% and $n$ subtrees $t_1$,\ldots, $t_n$.
%
Let us assume a commutative semiring~$\Semiring$ 
and a label theory~$\bar{\Phi}$ over~$\Semiring$ 
containing one set~$\Phi_{\Omega_k}$ for each $k \in [0..M]$.

\renewcommand{\call}[1]{\ensuremath \langle_{#1}}
\renewcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$

Let $\hat\Omega$ be the countable (unranked) alphabet obtained
from $\Omega$ by: 
$\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, with
$\Omegai = \Omega_0$, 
$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
$\Omegar = \{ \; \return{a} \mid a \in \Omega_{>0} \}$.
We associate to $\hat\Omega$
a label theory $\hat{\Phi}$ 
like in Section~\ref{se:SWVPA-def}.

\noindent
We define a linearization of trees of $\T(\Omega)$ into 
words of $\hat{\Omega}^*$ as follows:
\begin{description}
\item $\lin(a) = a$ for all $a \in \Omega_0$, 
\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) = 
       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$ 
       when $b \in \Omega_k$, $1 \leq k \leq M$.
\end{description}

% A \emph{regular tree grammar} over $\Omega$ 
% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
% $N$ is a finite set of non-terminal symbols denoted $q$..., 
% $q_\mathsf{i} \in N$ is the starting non-terminal, 
% $R$ is a finite set of production rules of the form
% $q_0 \to a(q_1\ldots q_k)$ where 
% $q_0, q_1, \ldots, q_k \in N$
% $a \in \Omega_k$.
% A tree $t \in \T(\Omega)$ is in the language of $G$ 
% if it can be generated from $q_\mathsf{i}$ by 
% non terminal replacement following the rules of $R$.
%
\begin{definition}  \label{def:SWTA}
A \emph{symbolic-weighted tree automaton} (\SWTA)
over $\Omega$, $\Semiring$, and $\bar\Phi$
is a triplet $A = \< Q, \init, \bar{\wei} >$ where
$Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
and $\bar{\wei}$ is a tuplet of transition functions containing
%$\wei_\varepsilon$ from $Q \times Q$ into $\Semiring$, and, 
$\wei_{k}: Q \times Q^{k} \to \Phi_{\Omega_k}$
for each $k \in [0..M]$. 
\end{definition}
%
%Like in Section~\ref{sec:SWAdef}, 
We define from $\bar{\wei}$ 
a transition function~$\wei$, %from $\bar\wei$
from 
$Q \times \Omega \times \bigcup_{k=0}^{M} Q^k$
into~$\Semiring$ by: %also called $\wei$ for simplicity, 
%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
\[
\begin{array}{rcll}
%\wei(q_0, \varepsilon, q_1) & = &  \wei_\varepsilon(q_0, q_1),\\ %\phi_\varepsilon\\
\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
\quad\mathrm{where~} \phi_{k} = \wei_{k}(q_0, q_1\ldots q_k).\\
\end{array}      
\]
Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
of a regular tree grammar~\cite{tata}, 
that replaces the non-terminal symbol $q_0$ by $a(q_1, \ldots, q_k)$. 
%in a step of tree building.
%
Such a grammar computes the weights of the derivation trees 
of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.

\noindent
The \SWTA of Definition~\ref{def:SWTA} defines a mapping 
from trees of $\T(\Omega)$ into the weights of~$\Semiring$,
based on the intermediate function~$\weight_A$
defined as follows for $q_0 \in Q$ and 
$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
with $0 \leq k \leq M$:
\begin{equation}
\begin{array}{rccl}
\weight_A(q_0, t) & = & 
%   \displaystyle\bigoplus_{q_1 \in Q} &
%   \wei(q, \varepsilon, q_1) \otimes \weight_A(q_1, t)\\
 \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
              \wei(q_0, b, q_1 \ldots q_k ) 
   \otimes \displaystyle\bigotimes_{i=1}^{k}
           \weight_A(q_i, t_i)
\end{array}
\end{equation}

@@

\begin{lemma}\label{lem:SWTA}
For all \SWTA $A$ over~$\Omega$, $\Semiring$ commutative, and $\Phi$,
there exists an effectively constructible \SWVPA $A'$ over 
$\hat\Omega$, $\Semiring$ and $\hat\Phi$ 
such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
\end{lemma} 
% 
\begin{proof}
Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is presented as above by a function
    %
We build 
$A' = \< Q', P', \init', \bar{\wei}', \final' >$,
%computing over $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$,
where $Q' = \bigcup_{k=0}^{M} Q^k$ is the set of sequences of state symbols of $A$, 
of length at most $M$, including the empty sequence denoted by $\varepsilon$, 
and where $P' = Q'$ and $\bar\wei$ is defined by:

@@

\section{Symbolic Weighted Parsing}
\label{sec:parsing}
Let us now apply the models and results of the previous sections %in order to define 
to the problem of parsing over infinite alphabet. %appropriate
%
Besides considering infinitely many possible of input symbols, 
handled with suitable language formalisms, 
this approach extends conventional parsing and weighted parsing
by computing a derivation tree modulo 
a generic distance between words,  
defined by a SW transducer given in input.
This enables considering finer word relationships than strict equality 
as in the conventional parsing approach, 
opening possibilities of quantitative analysis via this method.

%\subsection{Definition}
%
Let $\Sigma$ be a countable input alphabet 
and let $\Omega$ be a countable output ranked alphabet, 
with maximal rank value $M$, 
and $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$
be the alphabet with nesting symbols, associated like Section~\ref{sec:trees}, 
for the linearization of trees of $\T(\Omega)$
(remember that $\Omegai = \Omega_0$).
%
Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, bounded, complete 
and total semiring 
and let $\bar\Phi$ be a label theory over $\Semiring$
containing
$\Phi_\Sigma$, $\Phi_{\Sigma, \Omegai}$, as well as
$\Phii$, $\Phic$, $\Phir$, $\Phicr$
(following the notations of Section~\ref{se:SWVPA-def}).
It is moreover assumed computable and $k$-convex for some fixed $k$.

\noindent
Let us assume given the following input:
\begin{itemize}
\item a \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
defining a measure %between words 
$T: \Sigma^* \times \Omegai^* \to \Semiring$,

\item a \SWTA $A$ over $\Omega$, $\Semiring$, and $\bar\Phi$,
defining a measure $A: \T(\Omega) \to \Semiring$,
%a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
%      $A : \Omega^* \to \Semiring$,
\item an input word $s \in \Sigma^*$.
\end{itemize}
%
As explained in Section~\ref{sec:trees}, 
$A = \< Q, \init, \bar{\wei} >$ can be seen as a weighted regular tree grammar, 
that generates (weighted) trees by replacement of a state symbol $q_0$ (non-terminal), 
by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
A replacement rule $q_0 \to a(q_1,\ldots, q_k)$, 
of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$ 
according to Definition~\ref{def:SWTA}, 
corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
with set non-terminal symbols $Q$ and set of terminal symbols $\Omega_0$.
This actually is a slight generalization of CFG since 
each such production rule is labelled by  a symbol of $\Omega_{>0}$, 
hence parse trees %derivation trees
are trees of $\T(\Omega)$.
%
Another (more original) generalization is that the set of terminal symbols 
$\Omega_0$ may be infinite.

@@

\noindent 
\emph{Symbolic weighted parsing} is the problem, 
given the above input, 
to find a tree $t \in \T(\Omega)$ %nested word $t \in \Omega$ 
minimizing \( d(s, t) \otimes A(t)\)
\wrt $\leq_\oplus$, 
\ie such that: %Hence, it is the problem of finding 
%
\begin{equation}\label{eq:distance-lang}
d(s, t) \otimes A(t) = \displaystyle\bigoplus_{t' \in \T(\Omega)} d(s, t') \otimes A(t') 
\end{equation}
%
The measure expressed in~\eqref{eq:distance-lang} is called
called edit-distance between $s$ and $A$ in~\cite{Mohri03EDWA}.
%
%The input language can also be expressed as a \SWTA, or, 
%as a particular case, as a weighted context-free grammar, 
%converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
%
The problem of searching, in a WTA language, 
the best parse tree matching a given input,  
sometimes referred as~\emph{weighted parsing}
corresponds to SW parsing in the case of finite alphabets
and when the transducer $T$ characterizes identity
see \eg~\cite{Goodman99SemiringParsing} 
and~\cite{MorbitzVogler19weighted-parsing} for a more general 
weighted parsing framework.
%
%Indeed, it corresponds to the case where $T$
%accepts only the pairs $\<s, t>$ such that 
%$s$ is the projection of $t$ on $\Omegai$. 
%This can be done with a single state $q$ and 
%with transition rules of the form:
%\begin{description}
%\item[] $\wei(q, \varepsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
%\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
%\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
%\end{description}

\subsection{Computation}

\begin{proposition}
The problem of Symbolic Weighted  parsing 
can be solved in PTIME in the size of the input \SWT $T$, \SWTA $A$ %\SWVPA (or ) 
and input word $s$, 
and the computation time of the functions of the label theory.
\end{proposition}
\begin{proof} (sketch)
We follow a \emph{Bar-Hillel} construction, also called parsing by intersection.

@@

\subsection{Application to Automated Music Transcription}
%\input{transcription}

\section*{Conclusion}
% summary 
We have introduced weighted language models (SW transducers and visibly pushdown automata)
computing over infinite alphabets, 
and applied them to the problem of parsing 
with infinitely many possible input symbols (typically timed events).
%handled with suitable language formalisms, 
%
This approach extends conventional parsing and weighted parsing
by computing a derivation tree modulo 
a generic distance between words,  
defined by a SW transducer given in input.
This enables considering finer word relationships than strict equality 
as in the conventional parsing approach, 
opening possibilities of quantitative analysis via this method.

% discussion
Ongoing and future work include

\noindent
-- other theoretical properties of SW models
\eg extention to $n$-best, 
to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
(instead of \emph{bounded}, which is equivalent to $0$-\emph{closed})

\noindent
-- room to improve complexity for best-search algorithm
... modular approach with oracles ...

>>>>

Let $\Omega$ be a countable ranked alphabet, such that 
every symbol $a \in \Omega$ has a rank 
$\rank(a) \in [0..M]$ where $M$ is a fixed natural number.
We denote by $\Omega_k$ the subset of all symbols $a$ of $\Omega$
with $\rank(a) = k$, where $0 \leq k \leq M$, 
and $\Omega_{>0} = \Omega \setminus \Omega_0$.
%
\noindent 
The free $\Omega$-algebra of finite, ordered, 
$\Omega$-labeled trees is denoted by $\T(\Omega)$.
It is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
and for all $1 \leq k \leq M$, all $a \in \Omega_k$, 
and all $t_1, \ldots, t_k \in \T(\Omega)$, $a(t_1, \ldots, t_k) \in \T(\Omega)$.
%
% tree = single node (leave) labeled with a symbol of $a \in \Omegai$
% (such a tree is simply denoted by $a$)
% or the composition, denoted by $b(t_1,\ldots, t_n$) of a node labeled with $b$
% and $n$ subtrees $t_1$,\ldots, $t_n$.
%
Let us assume a commutative semiring~$\Semiring$ 
and a label theory~$\bar{\Phi}$ over~$\Semiring$ 
containing one set~$\Phi_{\Omega_k}$ for each $k \in [0..M]$.
%
\renewcommand{\call}[1]{\ensuremath \langle_{#1}}
\renewcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$

% A \emph{regular tree grammar} over $\Omega$ 
% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
% $N$ is a finite set of non-terminal symbols denoted $q$..., 
% $q_\mathsf{i} \in N$ is the starting non-terminal, 
% $R$ is a finite set of production rules of the form
% $q_0 \to a(q_1\ldots q_k)$ where 
% $q_0, q_1, \ldots, q_k \in N$
% $a \in \Omega_k$.
% A tree $t \in \T(\Omega)$ is in the language of $G$ 
% if it can be generated from $q_\mathsf{i}$ by 
% non terminal replacement following the rules of $R$.
%
\begin{definition}  \label{def:SWTA}
A \emph{symbolic-weighted tree automaton} (\SWTA)
over $\Omega$, $\Semiring$, and $\bar\Phi$
is a triplet $A = \< Q, \init, \bar{\wei} >$ where
$Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
and $\bar{\wei}$ is a tuplet of transition functions containing
%$\wei_\varepsilon$ from $Q \times Q$ into $\Semiring$, and, 
one function $\wei_{k}: Q \times Q^{k} \to \Phi_{\Omega_k}$
for each $k \in [0..M]$. 
\end{definition}
%
%Like in Section~\ref{sec:SWAdef}, 
We define from $\bar{\wei}$ 
a transition function~$\wei$, %from $\bar\wei$
from 
$Q \times \Omega \times \bigcup_{k=0}^{M} Q^k$
into~$\Semiring$ by: %also called $\wei$ for simplicity, 
%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
\[
\begin{array}{rcll}
%\wei(q_0, \varepsilon, q_1) & = &  \wei_\varepsilon(q_0, q_1),\\ %\phi_\varepsilon\\
\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{k}(a) &
\quad\mathrm{where~} \phi_{k} = \wei_{k}(q_0, q_1\ldots q_k).\\
\end{array}      
\]
%
\noindent
The \SWTA of Definition~\ref{def:SWTA} defines a mapping 
from trees of $\T(\Omega)$ into the weight values in~$\Semiring$,
based on the intermediate function~$\weight_A$
defined as follows for $q_0 \in Q$ and 
$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
with $0 \leq k \leq M$:
\begin{equation}
\begin{array}{rccl}
\weight_A(q_0, t) & = & 
%   \displaystyle\bigoplus_{q_1 \in Q} &
%   \wei(q, \varepsilon, q_1) \otimes \weight_A(q_1, t)\\
 \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
              \wei(q_0, b, q_1 \ldots q_k ) 
   \otimes \displaystyle\bigotimes_{i=1}^{k}
           \weight_A(q_i, t_i)
\end{array}
\end{equation}

@@

\noindent
Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
of a regular tree grammar $G$~\cite{tata}, 
that replaces the non-terminal symbol $q_0$ by $a(q_1, \ldots, q_k)$. 
%in a step of tree building.
%
%Such a grammar computes the weights of the derivation trees 
%of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
And the above production rule can also be seen as 
a rule of a weighted CF grammar, of the form
$[a]\, q_0 := q_1 \ldots q_k$ if $k > 0$,
and $q_0 := a$ if $k = 0$. 
In the first case, $a$ is a label of the rule, 
in the second case it is a terminal symbol.
On can observe that the weight of a labeled derivation tree $t$
of the weighted CF grammar associated to~$A$ as above, 
is $\weight_A(q, t)$, 
when $q$ is the start non-terminal.
%
We shall now establish a correspondence between such derivation tree $t$
and some word describing a linearization of $t$, 
in a way that $\weight_A(q, t)$ can be computed by a $\SWVPA$.

Let $\hat\Omega$ be the countable (unranked) alphabet obtained
from $\Omega$ by: 
$\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, with
$\Omegai = \Omega_0$, 
$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
$\Omegar = \{ \; \return{a} \mid a \in \Omega_{>0} \}$.

\noindent
We associate to $\hat\Omega$
a label theory $\hat{\Phi}$ 
like in Section~\ref{se:SWVPA-def}, 
%
\noindent
and we define a linearization of trees of $\T(\Omega)$ into 
words of $\hat{\Omega}^*$ as follows:
\begin{description}
\item $\lin(a) = a$ for all $a \in \Omega_0$, 
\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) = 
       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$ 
       when $b \in \Omega_k$ for $1 \leq k \leq M$.
\end{description}

\begin{proposition}\label{lem:SWTA}
For all \SWTA $A$ over~$\Omega$, $\Semiring$ commutative, and $\bar\Phi$,
there exists an effectively constructible \SWVPA $A'$ over 
$\hat\Omega$, $\Semiring$ and $\hat\Phi$ 
such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
\end{proposition} 
% 
\begin{proof}
Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is presented as above by a function
    %
We build 
$A' = \< Q', P', \init', \bar{\wei}', \final' >$,
%computing over $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$,
where $Q' = \bigcup_{k=0}^{M} Q^k$ is the set of sequences of state symbols of $A$, 
of length at most $M$, including the empty sequence denoted by $\varepsilon$, 
and where $P' = Q'$ and $\bar\wei$ is defined by:

@@

\section{Symbolic Weighted Parsing}
\label{sec:parsing}
Let us now apply the models and results of the previous sections %in order to define 
to the problem of parsing over infinite alphabet. %appropriate
%
%\subsection{Definition}
%
Let~$\Sigma$ be a countable input alphabet 
and~$\Omega$ be a countable output ranked alphabet, 
with maximal rank value~$M$, 
and let $\hat\Omega = \Omegai \uplus \Omegac \uplus \Omegar$
be the alphabet with nesting symbols associated to $\Omega$ like in Section~\ref{sec:trees},
for the linearization of trees of $\T(\Omega)$.
%(remember that $\Omegai = \Omega_0$).
%
Let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a 
commutative, bounded, and complete \marginpar{\tiny total?} semiring  
and let $\bar\Phi$ be an effective label theory over $\Semiring$,
containing $\Phi_\Sigma$, $\Phi_{\Sigma, \Omegai}$, as well as
$\Phii$, $\Phic$, $\Phir$, $\Phicr$
(following the notations of Section~\ref{se:SWVPA-def}).
%It is moreover assumed computable and $k$-convex for some fixed~$k$.
%
\noindent
We assume given the following input:
\begin{description}
\item[--] a \SWT $T$ over $\Sigma$, $\Omegai$, $\Semiring$, and $\bar\Phi$, 
defining a measure %between words 
$T: \Sigma^* \times \Omegai^* \to \Semiring$,

\item[--] a \SWTA $A$ over $\Omega$, $\Semiring$, and $\bar\Phi$,
defining a measure $A: \T(\Omega) \to \Semiring$,
%a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
%      $A : \Omega^* \to \Semiring$,
\item[--] an input word $s \in \Sigma^*$.
\end{description}
%
As explained in Section~\ref{sec:trees}, 
$A = \< Q, \init, \bar{\wei} >$ can be seen as a weighted regular tree grammar, 
or a labeled weighted CF grammar whose derivation trees are in $\T(\Omega)$, 
with weight defined by $A$.
%that generates (weighted) trees by replacement of a state symbol~$q_0$ (non-terminal), 
%by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
%A replacement rule $q_0 \to a(q_1,\ldots, q_k)$, 
%of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$ 
%according to Definition~\ref{def:SWTA}, 
%corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
%with set non-terminal symbols $Q$ and set of terminal symbols $\Omega_0$.
%
%This actually is a slight generalization of CFG since 
%each such production rule is labelled by  a symbol of $\Omega_{>0}$, 
%hence parse trees %derivation trees
%are trees of $\T(\Omega)$.
%
%Another (more original) generalization is that the set of terminal symbols 
%$\Omega_0$ may be infinite.

@@

\noindent 
\emph{Symbolic weighted parsing} is the problem, 
given the above input, 
to find a tree $t \in \T(\Omega)$ %nested word $t \in \Omega$ 
minimizing \( d(s, t) \otimes A(t)\)
\wrt $\leq_\oplus$, 
\ie such that: %Hence, it is the problem of finding 
%
\begin{equation}\label{eq:distance-lang}
d(s, t) \otimes A(t) = \displaystyle\bigoplus_{t' \in \T(\Omega)} d(s, t') \otimes A(t') 
\end{equation}
%
The measure expressed in~\eqref{eq:distance-lang} is called
called the edit-distance between~$s$ and~$A$ in~\cite{Mohri03EDWA}.
%
%The input language can also be expressed as a \SWTA, or, 
%as a particular case, as a weighted context-free grammar, 
%converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
%
The problem of searching, in a WTA language, 
the best parse tree matching a given input,  
sometimes referred as~\emph{weighted parsing},
corresponds to SW parsing in the case of finite alphabets
and when the transducer $T$ characterizes identity
see \eg~\cite{Goodman99SemiringParsing} 
and~\cite{MorbitzVogler19weighted-parsing} for a more general 
weighted parsing framework.
%
%Indeed, it corresponds to the case where $T$
%accepts only the pairs $\<s, t>$ such that 
%$s$ is the projection of $t$ on $\Omegai$. 
%This can be done with a single state $q$ and 
%with transition rules of the form:
%\begin{description}
%\item[] $\wei(q, \varepsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
%\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
%\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
%\end{description}

%\subsection{Computation}
%
\begin{proposition}
The problem of Symbolic Weighted  parsing 
can be solved in PTIME in the size of the input \SWT $T$, \SWTA $A$ %\SWVPA (or ) 
and input word $s$, 
and the computation time of the functions of the label theory.
\end{proposition}
\begin{proof} (sketch)
We follow a \emph{Bar-Hillel} construction, also called parsing by intersection.

@@

\paragraph{Application to Automated Music Transcription.}
%\input{transcription}
...

\section*{Conclusion}
% summary 
We have introduced weighted language models (SW transducers and visibly pushdown automata)
computing over infinite alphabets, 
and applied them to the problem of parsing 
with infinitely many possible input symbols (typically timed events).
%handled with suitable language formalisms, 
%
This approach extends conventional parsing and weighted parsing
by computing a derivation tree modulo 
a generic distance between words,  
defined by a SW transducer given in input.
This enables to consider finer word relationships than strict equality, 
%as in the conventional parsing approach, 
opening possibilities of quantitative analysis via this method.

% discussion
\noindent
Ongoing and future work include

\noindent
-- The study of other theoretical properties of SW models, 
such as the extension of the best search algorithm from $1$-best to $n$-best~\cite{Huang05kbest}, 
and to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
(instead of \emph{bounded}, which corresponds to $0$-\emph{closed}).

\noindent
-- ...there is room to improve the complexity bounds for the algorithms
... modular approach with oracles ...
